Laplacian envelope

### 2018-01-20 Notes/laplace.tex
chernoff etc

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
%
\input{sl_graphics_preamble.tex}
\graphicspath{{Figures/}}
%
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%prob distribution
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thb}{\boldsymbol{\theta}}
%synapse models
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{K}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%with activity indep processes
\renewcommand{\hom}{\mathbf{H}}
\newcommand{\Mh}{\widetilde{\M}}
\newcommand{\frgh}{\widetilde{\W}^{\mathrm{F}}}
\newcommand{\ench}{\widetilde{\enc}}
%eigenvectors
\newcommand{\evr}{\mathbf{u}}
\newcommand{\evl}{\boldsymbol{\eta}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{ideal}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\rh}{\hat{r}}
\newcommand{\CI}{\mathcal{I}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
%avoid double superscripts
\renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}

@@

We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold.
For large $N$, this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \operatorname{NNR} } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \operatorname{NNR} &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.

@@

As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRdef}
  \snr(t) = \frac{\av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}}
     {\sqrt{\var(\syn_\text{ideal}\cdt\syn(\infty))}},
\end{equation}
%
Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
If we also average over the time between storage and recognition, we have
%
\begin{equation}\label{eq:snrb}
  \snrb(\tau) = \intd[_0^\infty]{t} P(t\vert\tau) \snr(t)
   = \frac{1}{\tau} \intd[_0^\infty]{t} \e^{-t/\tau} \snr(t),
\end{equation}
%
This is similar to the average SNR up to time $\tau$, but with the hard cut-off replaced with a smooth exponential.
This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
%
\begin{equation}\label{eq:laplacedef}
  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t).
  \qquad
  \snrb(\tau) = \frac{A(1/\tau)}{\tau},
\end{equation}
%

First, let's look at the variances, remembering that the states and plasticity events of each synapse are independent and identically distributed:
%
\begin{equation}\label{eq:noise}
\begin{aligned}
  \var(\syn_\text{ideal}\cdt\syn(t))
    &= \sum_{\alpha\beta} \av{\syn^\alpha_\text{ideal}\syn^\alpha(t) \syn^\beta_\text{ideal}\syn^\beta(t)}
    - \prn{\sum_\alpha \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}}^2 \\
    &= \sum_{\alpha} \av{(\syn^\alpha_\text{ideal})^2(\syn^\alpha(t))^2}
    + \sum_{\alpha\neq\beta} \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}\av{\smash{\syn^\beta}_\text{ideal}\smash{\syn^\beta}(t)}
    \\&\phantom{= \sum_{\alpha}  \av{(\syn^\alpha_\text{ideal})^2(\syn^\alpha(t))^2}}
    - \prn{\sum_\alpha \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}}^2  \\
    &= N\av{1}
    + N(N-1)\av{\syn^1_\text{ideal}\syn^1(t)}^2
    - N^2\av{\syn^1_\text{ideal}\syn^1(t)}^2 \\
    &= N(1-\av{\syn^1_\text{ideal}\syn^1(t)}^2),
\end{aligned}
\end{equation}
%
where we used $\syn^\alpha=\pm1$.
We can compute $\var(\syn_\text{ideal}\cdt\syn(\infty))$ by taking $t\to\infty$.

@@

To write down the Hessian, we will first define some notation.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{aligned}
  R_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{[\fund(0)\enc\fund(s)]_{ik} - [\fund(0)\enc\fund(s)]_{jk}} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{aligned}
\end{equation}
%
Then we can write the hessian as
%
\begin{equation}\label{eq:hessian}
\begin{aligned}
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\pot)^2 (R_{ijkl} + R\trans_{ijkl})
     + f\pot(S_{ijkl} + S\trans_{ijkl}), \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
     - f\pot S_{ijkl} + f\dep S\trans_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
     - f\pot S\trans_{ijkl} + f\dep S_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\dep)^2 (R_{ijkl} + R\trans_{ijkl})
     - f\dep(S_{ijkl} + S\trans_{ijkl}), \\
\end{aligned}
\end{equation}
%
This is useful in numerical maximisation.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
%
\input{sl_graphics_preamble.tex}
\graphicspath{{Figures/}}
%
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%prob distribution
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thb}{\boldsymbol{\theta}}
%synapse models
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{K}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%with activity indep processes
\renewcommand{\hom}{\mathbf{H}}
\newcommand{\Mh}{\widetilde{\M}}
\newcommand{\frgh}{\widetilde{\W}^{\mathrm{F}}}
\newcommand{\ench}{\widetilde{\enc}}
%eigenvectors
\newcommand{\evr}{\mathbf{u}}
\newcommand{\evl}{\boldsymbol{\eta}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\rh}{\hat{r}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
%avoid double superscripts
\renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}

@@

We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold.
For large $N$, this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.

There are other measures of memory one could consider, \eg asymptotic error exponents.
For the Neyman-Pearson approach to hypothesis testing, the error exponent is the KL divergence:
%
\begin{equation*}
  D_\text{KL}\prn{P_0 \middle\Vert P_1} = \ln\nnr + \frac{1 + \snr^2}{2\nnr^2} - \frac{1}{2}.
\end{equation*}
%
For Bayesian hypothesis testing, the error exponent is the Chernoff distance:
%
\begin{equation*}
  D^*\prn{P_0 \middle\Vert P_1} = \max_\alpha \brc{
    \frac{1}{2} \ln\brk{\frac{1 + \alpha(\nnr^2-1)}{\nnr^{2\alpha}}} + \frac{\alpha(1-\alpha)\snr^2}{2[1 + \alpha(\nnr^2-1)]}
  }.
\end{equation*}
%

@@

As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRdef}
  \snr(t) = \frac{\av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}}
     {\sqrt{\var(\syn_\text{ideal}\cdt\syn(\infty))}},
\end{equation}
%
Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
If we also average over the time between storage and recognition, we have
%
\begin{equation}\label{eq:snrb}
  \snrb(\tau) = \intd[_0^\infty]{t} P(t\vert\tau) \snr(t)
   = \frac{1}{\tau} \intd[_0^\infty]{t} \e^{-t/\tau} \snr(t),
\end{equation}
%
This is similar to the average SNR up to time $\tau$, but with the hard cut-off replaced with a smooth exponential.
This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
%
\begin{equation}\label{eq:laplacedef}
  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t),
  \qquad
  \snrb(\tau) = \frac{A(1/\tau)}{\tau}.
\end{equation}
%

First, let's look at the variances, remembering that the states and plasticity events of each synapse are independent and identically distributed:
%
\begin{equation}\label{eq:noise}
\begin{aligned}
  \var(\synid\cdt\syn(t))
    &= \sum_{\alpha\beta} \av{\synid^\alpha \syn^\alpha(t) \synid^\beta \syn^\beta(t)}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2 \\
    &= \sum_{\alpha} \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}
    + \sum_{\alpha\neq\beta} \av{\synid^\alpha\syn^\alpha(t)}\av{\smash{\synid^\beta}\smash{\syn^\beta}(t)}
    %\\&\phantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2  \\
    &= N\av{1}
    + N(N-1)\av{\synid^1 \syn^1(t)}^2
    - N^2\av{\synid^1 \syn^1(t)}^2 \\
    &= N(1-\av{\synid^1 \syn^1(t)}^2),
\end{aligned}
\end{equation}
%
where we used $\syn^\alpha=\pm1$.
We can compute $\var(\syn_\text{ideal}\cdt\syn(\infty))$ by taking $t\to\infty$.

@@

To write down the Hessian, we will first define some notation.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{aligned}
  R_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{[\fund(0)\enc\fund(s)]_{ik} - [\fund(0)\enc\fund(s)]_{jk}} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{aligned}
\end{equation}
%
Then we can write the hessian as
%
\begin{equation}\label{eq:hessian}
\begin{aligned}
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\pot)^2 (R_{ijkl} + R\trans_{ijkl})
     + f\pot(S_{ijkl} + S\trans_{ijkl}), \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
     - f\pot S_{ijkl} + f\dep S\trans_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
     - f\pot S\trans_{ijkl} + f\dep S_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
     (f\dep)^2 (R_{ijkl} + R\trans_{ijkl})
     - f\dep(S_{ijkl} + S\trans_{ijkl}), \\
\end{aligned}
\end{equation}
%
This is useful in numerical maximisation.
